---
title: "HOWARD_Skills_Test_3_Markdown"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exam 3

First I will load some packages that will be needed.
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(janitor)
library(broom)
```

For the first task, I'm going to recreate the first plot shown and save it.

```{r message=FALSE, warning=FALSE}
sal <- read.csv("./FacultySalaries_1995.csv")
names(sal)
names(sal)

fult <- sal %>% 
  select(c("FedID", "UnivName", "State", "Tier", "NumInstructors", "AvgProfSalaryAll", "AvgProfCompAll", "NumFacultyAll", contains(c("Full")))) %>% 
mutate(Rank = "Full") %>% 
  rename(AvgSalary = AvgFullProfSalary) %>% 
  rename(AvgComp = AvgFullProfComp) %>% 
rename(NumProfs = NumFullProfs)

asso <- sal %>% 
  select(c("FedID", "UnivName", "State", "Tier", "NumInstructors", "AvgProfSalaryAll", "AvgProfCompAll", "NumFacultyAll", contains(c("Assoc")))) %>% 
  mutate(Rank = "Assoc") %>% 
  rename(AvgSalary = AvgAssocProfSalary) %>% 
  rename(AvgComp = AvgAssocProfComp) %>% 
  rename(NumProfs = NumAssocProfs)

names(asso)
assi <- sal %>% 
  select(c("FedID", "UnivName", "State", "Tier", "NumInstructors", "AvgProfSalaryAll", "AvgProfCompAll", "NumFacultyAll", contains(c("Assist")))) %>% 
  mutate(Rank = "Assist") %>% 
  rename(AvgSalary = AvgAssistProfSalary) %>% 
  rename(AvgComp = AvgAssistProfComp) %>% 
  rename(NumProfs = NumAssistProfs)

clean <- rbind(fult, assi, asso)

cclean <- clean[clean$Tier != "VIIB" ,]

# Recreate Figure 1
ggplot(cclean, aes(x=Rank, y=AvgSalary, fill = Rank)) +
  geom_boxplot() +
  facet_wrap(~Tier) +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  labs(y="Salary", 
       x="Rank")

# Save the graph
jpeg("./HOWARD_Fig_1.jpg")
ggplot(cclean, aes(x=Rank, y=AvgSalary, fill = Rank)) +
  geom_boxplot() +
  facet_wrap(~Tier) +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  labs(y="Salary", 
       x="Rank")
dev.off()
```
For task 2, I need to export an ANOVA table.
```{r message=FALSE, warning=FALSE}
mod1 <- aov(data = cclean,
            formula = AvgSalary ~ State + Tier + Rank)
anv <- summary(mod1)

capture.output(anv, file = "./Salary_ANOVA_Summary.txt")
```
For Task 3, I will load a new dataset and clean it.
```{r message=FALSE, warning=FALSE}
oil <- read.csv("./Juniper_Oils.csv")

coil <- pivot_longer(oil,
                     c("alpha.pinene","para.cymene","alpha.terpineol","cedr.9.ene","alpha.cedrene","beta.cedrene","cis.thujopsene","alpha.himachalene","beta.chamigrene","cuparene","compound.1","alpha.chamigrene","widdrol","cedrol","beta.acorenol","alpha.acorenol","gamma.eudesmol","beta.eudesmol","alpha.eudesmol","cedr.8.en.13.ol","cedr.8.en.15.ol","compound.2","thujopsenal"),
                     names_to = "Compound",
                     values_to = "Concentration")
```
Now I will attempt to recreate Figure 2.
```{r message=FALSE, warning=FALSE}
ggplot(coil, aes(x=YearsSinceBurn, y=Concentration)) +
  geom_smooth() +
  theme_minimal() +
  facet_wrap(Compound ~ ., scales="free")

# Save the graph
jpeg("./HOWARD_Fig_2.jpg")
ggplot(coil, aes(x=YearsSinceBurn, y=Concentration)) +
  geom_smooth() +
  theme_minimal() +
  facet_wrap(Compound ~ ., scales="free")
dev.off()
```
For Task 5, I need to show which compounds are significant to the Concentration value using a generalized linear model.
```{r message=FALSE, warning=FALSE}
mod2 <- glm(data = coil,
            formula = Concentration ~ Compound * YearsSinceBurn)
summary(mod2)

tidy(mod2) %>% 
  filter(p.value < .05)
```
End of Skills Test 3.